---
title: "Simplest path to calculating biomass using census data and allodb"
author: "Mauro Lepore"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
# This chunk named setup will run before any other code (https://goo.gl/BeM2Uu)
set.seed(1014)
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  # Figures (http://r4ds.had.co.nz/)
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = 0.618,
  fig.show = "hold"
)
```

```{r packages, message=FALSE}
library(tidyverse)
library(bmss)
library(allodb)
```

Maybe add some of these helpers to __allodb__, __bmss__, or elsewhere.

```{r helpers}
# List all datasets of a package.
datasets <- function(package) {
  sort(utils::data(package = package)$results[ , "Item"])
}

compare_names <- function(x, y) {
  list(x, y) %>% 
    purrr::map(names) %>% 
    purrr::reduce(setdiff)
}

add_species <- function(.census, .species) {
  suppressMessages({
    .species %>% 
      tidyr::unite("species", Genus, Species, sep = " ") %>% 
      dplyr::select(species, sp) %>% 
      dplyr::right_join(.census)
  })
}

add_site <- function(.census, site) {
  found_site <- grep(site, allodb::sites_info$site, ignore.case = TRUE, value = TRUE)
  if (identical(length(found_site), 0)) {
    rlang::abort( glue::glue("Can't find any site mathing {site}."))
  }
  
  rlang::inform(glue::glue("Using site {rlang::expr_label(found_site)}."))
  tibble::add_column(.census, site = found_site)
}
```

### Database data

All datasets.

```{r}
datasets("allodb")
```

The core table of __allodb__ is `equations`.

```{r}
glimpse(equations)
```

Some equations need to be fixed. Here we'll use the good ones.

```{r}
drop_bad_equations(equations)
```

`master()` combines multiple database tables. Glimpse excluding problematic equations.

```{r}
drop_bad_equations(master())
```

### Census data

```{r}
scbi_tree1
```

This dataset is not structured exactly as a reference dataset from Luquillo but the difference is irrelevant to calculating biomass. The column `agb` is not useful and `DBHID` is equivalent to `MeasureID`.

```{r}
reference_tree <- fgeo.data::luquillo_tree5_random

# In scbi but not luquillo
compare_names(scbi_tree1, reference_tree)
# In luquillo but not scbi
compare_names(reference_tree, scbi_tree1)
```

Let's add columns `site` and `species`.

FIXME: Wrap into a single `add_site_sp()` function, or `prepare_census()`, or `as_allodb_census()` (with corresponding `new_allodb_census()`, and `validate_allodb_census()` -- which is responsible to check stuff that then no longer needs to be checked downstream in __bmss__)?

```{r}
scbi <- scbi_tree1 %>% 
  add_species(scbi_species) %>% 
  add_site(site = "SCBI")

scbi %>% select(site, species, everything())
```

* FIXME: Update `bmss::get_allometry()` to no longer default to use the dummy dataset build in __bmss__ but `allodb::equation` (or `allodb::master()` or similar).

* FIXME: Rename as needed internals to work with new names (`old` = `new`):
   * `eqn` = `equation_allometry`.
   * `eqn_source` = `eqn_source`.
   * `eqn_type` = `allometry_specificity`.

```{r}
# FIXME
get_allometry(scbi, "site", "species", "dbh")
```

* FIXME: Update __bmss__ to work with one category, say, "Family", or "Species". Then further update to deal with all other categories.

```{r}
master()$allometry_specificity %>% unique()
```

