---
title: "Integrating bmss and allodb"
author: "maurolepore@gmail.com"
date: "2018-03-19"
output:
  rmarkdown::html_document:
    theme: united
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
set.seed(1014)

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold")
```

My goal is to integrate __allodb__ -- at least a simple subset of it -- with __bmss__. This document has two sections:

* In the first section I explore the `equations` dataset; I realize that some parameters are missing. This helps me to ask Erika the right question: Where can I find the missing parameters? The answer is, "In the `allodb_master` dataset.

* In the second section I try to integrate the parameters into the `equation` column.

# Section 1

In this section my goal is to explain what columns are missing from the current `equations` dataset. I'll show two examples, starting with a simplistic one.

```{r pkg}
library(tidyverse)
library(bmss)
library(allodb)
library(here)
```

## 1. Simplistic example

### Revisiting `biomass_per_indiv()`

Let's start with a little reminder of what the current code of __bmss__ does. 

This is a simplistic model of what the equations dataset might look like.

```{r}
eqn <- bmss::toy_default_eqn

eqn
```

Now say that we have a census dataset like this one:

```{r}
cns <- bmss::user_data

cns
```

We can calculate biomass with our currently naive `biomass_per_ind()` function.

```{r}
biomass_per_ind(cns, site = "site", sp = "sp", dbh = "dbh")
```

### Braking down `biomass_per_indiv()`

Let's brake down what `biomass_per_ind()` does to better understand what columns we need from the `equations` dataset.

* Step 1: Join the census data with the equations data.

```{r}
joint <- left_join(cns, eqn)

joint
```

To make things simpler, let's keep only the columns that matter.

```{r}
joint <- select(joint, dbh, eqn)

joint
```

* Step 2: Replace the string "dbh" of the column `eqn` with the corresponding value from the column `dbh`.

```{r}
joint <- mutate(
  joint, 
  eqn_replaced = str_replace(eqn, "dbh", as.character(dbh))
)

joint
```

* Step 3: Evaluate the equation strings in `eqn_replaced` to calculate biomass (`bmss`).

```{r}
dplyr::mutate(
  joint, 
  bmss = map_dbl(.data$eqn_replaced, ~eval(parse(text = .x), envir = joint))
)
```

## 2. An example that is a bit more realistic

The equations in the `equations` dataset have many parameters -- not only `dbh`. And we need one column for each of those parameters.

Let's glimpse the dataset, and focus on the `equation` column.

```{r}
glimpse(allodb::equations)
```

The column `equation` has many parameters, such as `a`, `b` and `c`. To show how the code may deal with these parameters, I'll create a small dataset with only the parameter `a`.

```{r}
joint2 <- tibble::tribble(
            ~eqn, ~a, ~dbh,
  "33 * dbh * a",  5,   10,
  "24 * dbh * a",  1,   15,
  "12 * dbh * a",  2,   23
)

joint2
```

And I will assume that this dataset is already merged with the user data, so it already has a `dbh` column -- that is, here I start at the step 2 above.

* Step 2: 
    * Replace the string "dbh" of the column `eqn` with the corresponding value from the column `dbh`.
    * Same with `a`.

```{r}
joint2 <- mutate(
  joint2, 
  eqn_replaced = str_replace(eqn, "dbh", as.character(dbh)),
  eqn_replaced = str_replace(eqn_replaced, "a", as.character(a))
)

joint2
```

* Step 3: Evaluate the equation strings in `eqn_replaced` to calculate biomass (`bmss`).

```{r}
dplyr::mutate(
  joint2, 
  bmss = map_dbl(.data$eqn_replaced, ~eval(parse(text = .x), envir = joint2))
)
```



# Section 2

In this section I show that we can programatically replace the parameters `a`-`d` from `equations$equation`. Also I identify a few parameters -- other than `a`-`d` and `DBH` -- that are missing and we'll need to somehow feed into the code for it to to its job.

Get data.

```{r, results="hide"}
master <- readr::read_csv(
  here::here("data-raw/allodb_master.csv"),
  col_types = allodb::type_allodb_master()
)

# Tidy
master <- master %>% 
  # Add spacing for readability (http://style.tidyverse.org/syntax.html#spacing)
  mutate(
    equation = formatR::tidy_source(text = equation)[[1]]
  ) %>% 
  # Show most key column first
  select(equation, everything())
```

```{r}
master
```


Explore distinct equations

```{r}
master %>%
  select(equation) %>% 
  distinct() %>% 
  print(n = nrow(.))
```

Notice `NA`s. What should we do with this?

```{r}
is_missing <- is.na(master$equation) | grepl("NA", master$equation)
master %>% 
  # Find both: missing value `NA` and the literal string "NA"
  filter(is_missing)
```

In the meantime, I keep only non missing equations.

```{r}
non_missing <- master %>% filter(!is_missing)
```

Replacing a-d.

```{r}
replaced <- non_missing %>%
  mutate(
    eqn_replaced = str_replace_all(equation, "a", as.character(a)),
  ) %>% 
  mutate(
    eqn_replaced = str_replace_all(eqn_replaced, "b", as.character(b)),
    eqn_replaced = str_replace_all(eqn_replaced, "c", as.character(c)),
    eqn_replaced = str_replace_all(eqn_replaced, "d", as.character(d)),
  ) %>% 
  select(eqn_replaced)

replaced
```

